\documentclass[11pt]{article}
\usepackage{enumerate}
\usepackage[T1]{fontenc}
\usepackage[top=1in, bottom=1in, left=1.25in, right=1.25in]{geometry}
\newcommand{\itab}[1]{\hspace{0em}\rlap{#1}}
\newcommand{\tab}[1]{\hspace{1em}\rlap{#1}}
\title{Homework 1st for Computational Complextiy}
\author{Shenlong Gu}
\date{sg3301@columbia.edu, 14, September 2015}
\begin{document}
\maketitle
\part{}
\def \mf {$M_{1}$}
\def \ms {$M_{2}$}
\def \mt {$M^{'}$}
\def \Lt {$L^{'}$}
We make some assumptions, turning machine \mf{} decides language $L$ and turning machine \ms{} decides language \Lt{}  \\
\begin{enumerate}[a)]
\item  {\bf union} \\
        We give a turning machine \mt, which works as follows. \\
        For a given string $l$,\\ 
        Simulate $l$ on \mf{}, \\ 
        if \mf{} accepts $l$: \\ 
        \itab{} \tab{\mt{} accepts $l$.} \\ 
        else: \\
        \itab{} \tab{simulate $l$ on \ms{}.} \\
        \itab{} \tab{if \ms{} accepts $l$:} \\
        \itab{} \tab{} \tab{\mt{} accepts $l$.} \\
        \itab{} \tab{else} \\
        \itab{} \tab{} \tab{\mt{} rejects $l$.} \\ 
        We can find \mt{} decides the language union of $L$ and $L^{'}$, so union operation is closed for two decidable languages.

\item {\bf concatenation} \\ 
        We give a turning machine \mt{}, which works as follows. \\
		For a give string $l$ with length n. \\
		We enumerates $i$ from 0 -> $n$ to generate $n + 1$ pairs $(i, n - i)$ \\
        For each pair$(i, n - i)$:\\ 
        \itab{} \tab{we simulate $l[0:i]$ on machine \mf{} and $l[i:n]$ on machine \ms{}} \\
        \itab{} \tab{if \mf{} accepts $l[0:i]$ and \ms{} accepts $l[i:n]$} \\
        \itab{} \tab{} \tab{\mt{} accepts $l$} \\
        \itab{} \tab{else} \\
        \itab{} \tab{} \tab{continue} \\
        End \\
        \mt{} rejects $l$. (because after enumerating each pair, we can not find a solution) \\ 
        We can find \mt{} decides the language concatenation of $L$ and \Lt{}, so concatenation operation is closed for two decidable languages. 

\item {\bf complementation} \\ 
        We give a turning machine \mt{}, which works as follows: \\
        For a given string $l$, \\ 
        simulate l on \mf{}.\\
        if \mf{} accepts $l$:\\
        \itab{} \tab{\mt{} rejects $l$.}\\
        else: \\
        \itab{} \tab{M' accepts $l$.} 
\item {\bf intersection} \\
        We give a turning machine \mt{}, which works as follows: \\
		for a given string $l$, \\ 
        simulate $l$ on \mf{}, \\ 
        if \mf{} rejects $l$: \\
        \itab{} \tab{\mt{} rejects $l$.} \\
		else: \\
        \itab{} \tab{simulate $l$ on \ms{},}\\
        \itab{} \tab{if \ms{} rejects $l$:} \\
        \itab{} \tab{} \tab{\mt{} rejects $l$.} \\
        \itab{} \tab{else:} \\ 
        \itab{} \tab{} \tab{\mt{} accepts $l$.} \\
        We can find \mt{} decides the language intersection of $L$ and \Lt{}, and intersection operation is closed for two decidable languages.
\end{enumerate}
\part{}
    We will prove for any lanauge which can be decided by an original turning machine can be decided by a new designed turning machine
    and for any language which can be decided by the new designed turning machine can be decided by an original turning machine.
\begin{enumerate}[a)]
\item {\bf First} \\
    For a language which can be decided by an orignal turning machine \mf{}, we can invents a new designed turning machine \ms{} which works as follows: \\
    We add one more character into the character sets to set it to the left of the start point of input( just like a left-end),
    then we use the same state change function as \mf{} does, (just the left end is replaced by the new character), obviously this new designed turning machine \ms{} decides
    the same language as \mf{} does.
\item {\bf Second} \\
    For a language which can be decided by a new designed turning machine \mf{}, we can design an original turning machine \ms{} as follows:  \\
    Three are conditions below. \\ 
    1.\mf{} scans in the range[left, right] \\ 
        \ms{} has the same state change as \mf{} does including state change and character write \\
    2.\mf{} moves from rightmost to the right of rightmost(which means go outside rightmost) \\
        \ms{} moves right to the right of rightmost of the input \\
    3.\mf{} moves left/right and in the range[right + 1, ++] \\
        everytime \mf{} moves one place left/right, \ms{} moves two places left/right, and do similar state change. \\ 
    4.\mf{} moves from the right + 1 position  left to the rightmost of the input \\
        \ms{} moves left one place to the rightmost of the input \\ 
    5.\mf{} moves from letmost to the left of leftmost(which means go outside leftmost) \\
        \ms{} moves to the rightmost of the input and moves two places right. \\
    6.\mf{} moves left/right and in the range[--, left - 1] \\
        everytime \mf{} moves one place left/right, \ms{} moves two places right/left, and do similar state change. \\ 
    7.\mf{} moves from the left - 1 position right to the leftmost of the input \\
        \ms{} moves to the leftmost of the input \\

    All steps above are just cursors change when an input runs in the \mf{} and \ms{}. \\
    In the actual operation, we can set one more characters to tag the rightmost, and the core idea is the following:
    First, We can see the difference is that the new-designed turning machine has two direction infinite tapes, however, original turning machine has only one.
    But we can extend the original turning machine to change it to "have two direction tapes", we use rightmost 1, 3, 5, 7, 9 ...2n + 1 positions to correspond to
    (count) the right 1,2,3...n positions, use rightmost 2, 4, 6, 8, .... 2n positions to correspond to the left, 1, 2, 3....n positions. And we do similar state changes
    in the \ms{} as \mf{} does, so we can see For a language which can be decided by a new designed turning machine \mf{}, we can design an original turning machine \ms{}.
    So we can see the power of these two turning machines is the same, can decide the same class of languages.
\end{enumerate}
\part{}
First, we will prove that a write-twice turning machine has the same power as the origin turning machine. For each of the write operation in the origin turning machine,
we copy the entire tape used right to a new unused portion and marked the tape copied(then we know the head of new tape because the left of it has been marked, and the whole used tape has been copied), so a cell in the tape needs only to be marked and writen (write-twice most)), for the position to be updated, we marked it first, and when copying, if it is already marked, we update the target cell, othervise copy the cell and mark this cell.
so a write-twice turning machine has the same power as the origin turning machine. \\
Second, we will prove write-once turning machine can work the same. To marked a cell, we split this cell into two parts, which means we use two cells, one to record character (can be update at most once), and a flag cell(marked or unmarked, can be updated only once), so this write-once turning machine works as the same logic as the write-twice turning machine (but each cell will be written at most once). \\
So in total, write-once turning machines can decide the same class of languages as the original turning machines.
\part{}
First, we can find this turning machine can do state-change and move left/right, no less powerful than DFA, so this turning machines can decide regular languages. \\
Next, we will prove this kind of modified turning machines can only decide regular languages. \\
First, even thought we can write on the rest blank position, we can not simply copy the whole string to the rest and do as original turning machines does.
Because we can not mark input string which means, we can not remember which position we are copying now. So extra space is useless. Next we will prove this 2-way DFA can not bring more power to this kind machine than DFA, can noly decide regular languages. \\
We will prove this by using Myhill-Nerode Theorem. First, we will see how this 2-way DFAs work.
we can split the input to $x, z$. At the beginning, the head is in the left of$x$, and within a series of operation, the machine will ends or the heads goes to z
with a state $q$, and within $z$ this machine will do a series of operation, and the machine will end or the head will go back to $x$ with a state $p$. This loop will continue until the machine ends.\\
We can see $x$ as a component, it will accept a state $p$ from right or start from left, and do a series of operation finally the machine will stop or will emit a state $q$ to the right.\\
So we can construct a state-change function: T$(p)$ = $q$ specified by input $x$, which means $x$ will accept from right by a state $p$ and emit a state $q$ to right, but there are spcial output, T$(p)$ = $accept$, T$(p)$ = $reject$ which means input a state $p$, the machine will reject or accept, there are also special input T$(start)$ = $q$, which means when the machine starts, if the head emits $x$, it will emit a state $q$. \\
Then we can easily if two inputs $x, y$ with same function T will be equivalent to each other, because for a input $z$ if $xz$ $\in$ $L$, $yz$ must $\in$ to $L$, because with the same state from $z$, $x, y$ will emit same state to $z$, all the machine will end with same results for $x$ or $y$. \\  
And it is clear, there are just finite function T. If the state number is $k$, the number of function T is almost $k^{k}$, a finite number. That's to say,
the number of equivalent classes for language $L$ is finite, and according to Myhill-Nerode Theorem, language $L$ is regular languages, so this modified turning machine can only decide regular languages. 
\part{}
\end{document}
