\documentclass[11pt]{article}
\usepackage{enumerate}
\usepackage[T1]{fontenc}
\usepackage[top=1in, bottom=1in, left=1.25in, right=1.25in]{geometry}
\newcommand{\itab}[1]{\hspace{0em}\rlap{#1}}
\newcommand{\tab}[1]{\hspace{1em}\rlap{#1}}
\title{Homework 2 for Intro to Computational Complextiy}
\author{Shenlong Gu}
\date{917-544-8927, sg3301@columbia.edu, 27, September 2015}
\begin{document}
\maketitle
\part{}
\def \vf {$V_{1}(x,c)$}
\def \vs {$V_{2}(x,c)$}
\def \v {$V(x,c)$}
\def \mf {$M_{1}$}
\def \ms {$M_{2}$}
\def \lf {$L_{1}$}
\def \ls {$L_{2}$}
    We assumes a Non-DTM \mf{} decides the language \lf, and 
    a Non-DTM \mf{} decides the language \ls. \\
    Let \vf{} be the verifier for \mf{} and \vs{} be the verifier
    for \ms{}.
\begin{enumerate}
\item {\bf union} \\
    We will construct a poly-time \v{} for the language, the union of 
    \lf{} and \ls{} to solve this problem.\\
    for any input $x$, given a stratage input string $c$, let \v{} = 1 iff \vf{} = 1 and \vs{} = 1. \\
    It is easy to see \v{} is the verifier which runs in polynomial time 
    given input $x$ and $c$ and \v{} is the verifier for the language, the
    union of \lf{} and \ls{}. So NP is closes under union operation. 
\item {\bf concatenation} \\
    We will construct a poly-time \v{} for the language, the union of 
    \lf{} and \ls{} to solve this problem.\\
    for any input $x$, the stratage input $c$ which is the union-encoding of the stratage $c_{1}$ for \vf{}, the stratage $c_{2}$ for \vs{}. \\
    \v{} works as the follows, creates $n + 1$ pairs, $(i, n - i), 0 \leq{} i \leq n$. Decode $c$ into $c_{1}$ and $c_{2}$ \\
    foreach pair $(i, n - i)$: \\
    \itab{} \tab{if {$V_{1}(x[0:i],c_{1})$} == 1 and {$V_{2}(x[i:n],c_{2})$} == 1:}\\
    \itab{} \tab{} \tab{\v{} = 1, accepts} \\
    end
    \v{} = 0, rejects. (Because no pairs work given the stratage $c_{1}$ 
    and $c_{2}$. \\
    It is easy to see \v{} is the verifier which runs in polynomial time 
    given input $x$ and $c$ and \v{} is the verifier for the language, concatenation of \lf{} and \ls{}. So NP is closes under concatenation operation. 
\item {\bf star} \\
    We will construct a poly-time \v{} for the language, the union of 
    \lf{} and \ls{} to solve this problem.\\
    Given an input $x$, we let $c$ to encode a segmentation of input $x$ into 
    $x_{1},x_{2},...x_{k}$, and a stratage of \vf{} to these $k$ segmentation $c_{1},c_{2},...c_{k}$. \\
    Then, \v{} works as follows: \\
    decode $c$ first, if $c$ can not be decoded, rejects it. \\
    for $i = 0$ to $k$: \\
    \itab{} \tab{if $V_{1}(x_{i},c_{i})$ == 1:} \\
    \itab{} \tab{} \tab{\v{} = 1, accepts} \\
    end \\
    \v{} = 0, rejects.\\
    It is easy to see \v{} is the verifier which runs in polynomial time 
    given input $x$ and $c$ and \v{} is the verifier for the language, star of \lf{}. So NP is closes under star operation. 
\end{enumerate} 
Then to prove $G$ and $H$ isomorphic. Given an input $x$ with length $n$, we encode a permutation of $n$ which is a node mapping from $G$ and $H$, 
we give a verifier \v{} given the input $x$ and encoding mapping $c$, it is easy to judge if two maps are isomorphic under this mapping in poly-time 
(just check if each mapping edge exists or doesn't exist in two graphs).
\part{}
    First, we will show that if a graph is bipartite, then it can not have a cycle containing an odd number of nodes. 
    Assume there is a such cycle the bipartite, $n_{1}$, $n_{2}$,.....$n_{2*k+1}$. We color the nodes into two part(blue and red)
    . if $n_{1}$ is red, $n_{2}$ is blue, then $n_{2*k+1}$ will be red the same as $n_{1}$, and there is an edge between $n_{1}$ and $n_{2*k+1}$,
    so there comes a contradictory. \\ \\
    Second, if there is no such cycle, we will show it is a bipartite. We also color the node, we randomly selects a node, and gives a color (for example, red),
    We traverse the edge from this node using dfs or bfs, everytime there is an edge between two nodes, we color the child node with a opposite color from its parent, when there is a back edge. Because there is no cycyle with an odd number of nodes, this node will not result in a contradictory with its ancestor, so we can color all nodes without contradictory, so it is a bipartite. \\ \\
    Third, we will show bipartite problem $\in$ NL. \\ 
    We change the bipartite judging problem into a existing problem. As proved above, we can change a bipartite judging problem by deciding that there is no cycle containing odd number nodes.\\
    Then we think about the complement problem(because NL equals CO-NL): how to decide that if there is a cycle containing odd number nodes. \\
    To use NDTM to solve this problem: \\
    First, we nondeterministicly selects a start node, $s$, and set a next node $t$ to nondeterministicly be a child of $s$, and set a counter to be zero .\\
    while counter $\leq$ $n$:\\
    ($n$ is total number of nodes, because the length of the cycle can not exceed $n$) \\
    \itab{} \tab{if $t$ equals $s$:} \\
    \tab{} \tab{} \tab{if counter is odd:}\\
    \tab{} \tab{} \tab{} \tab{accepts it}\\
    \tab{} \tab{} \tab{else:}\\
    \tab{} \tab{} \tab{} \tab{rejects it}\\
    \itab{} \tab{We nondeterministicly set $t$ to be a child of $t$,(if there is no child, we rejects)}\\
    \itab{} \tab{counter = counter + 1} \\ 
    We see that this NDTM decides the problem (odd-number-nodes cycle existing problem), and the space it uses is just start node number, next node number, counter which cost just logarithmic space. So bipartite judging problem $\in$ NL. 
\part{}
First we assume P = NP, if we can show NEXPTIME = EXPTIME, we get the origin problem proved. \\
Give a language $L$ in NTIME($2^{n^{k}}$), we will prove it in some EXPTIME. \\
We construct a $L^{'}$:pad(L, $2^{n^{k}}$), we can easily prove it in NP and because P = NP, so $L^{'}$ in poly-time, and because we can decide 
$L^{'}$ in P, we can decide $L$ in EXPTIME($2^{n^{k}}$), (just padding $L$ to be $L^{'}$ and simulate $L^{'}$ in poly-time which is some EXPTIME relative to $L$, so we can show NEXPTIME = EXPTIME, so the origin problem is proved.
\part{}
We first assume P = SPACE(n), and we can get a contradiction. \\
Given a language $L$ in SPACE($n^{2}$), we construct a language $L^{'}$: pad($L$, $n^{2}$). It is easy to show $L^{'}$ in SPACE(n), and because P = SPACE(n), we can solve $L^{'}$ in poly-time which means we can solve $L$ in poly-time (just use 0($n^{2}$) to pad, use 0($n^{2*c}$) to decide $L^{'}$), so we can find SPACE($n^{2}$) $\subset$ P, which means SPACE($n^{2}$) $\subset$ SPACE(n), which causes a contradiction by space hierarchy theorem. 
\part{}
First, we asssume a turning machine with a read-only input tape uses space $s(n)$ that is smaller than $c log log n$, then we show that it will uses constant space. \\
We consider a configuration worker head is in $j$ position, the status is $q$, the content in worker tape is $txt$, we will show for each input position $i$ in the input text, such configurations will at most have $nn$ = |Q| * ($|\sum|^{s(n)}$) * $s(n)$ = O($2^log(s(n))$) \\ 
Then we consider a sequence which means in the execution of $M$, the input header may run into position $i$ in the input tape for several times and we get a sequence of configurations above. We calculate for different sequences at position $i$ may have (we can not get into same configuration for twice, thus we will not halt), $\sum\limits_{i=1}^{nn} nn*(nn-1)*(nn-2)*...(nn-i+1)$ < O($nn^{nn}$) < O($2^{2^{2 * s(n)}}$) = O(n),(a linear bound). \\
O(n) means for a $n$ $\geq$ a given $n_{0}$, $s(n)$ <= $n / 3$. And we assume there exists a $n_{1}$ where $s(n_{1})$  $>$ $s(i)$ for $i$ $\leq$ $n_{1}$. Other wise we can easily conduct every $s(i)$ will equals $s(n_{0})$ which means space usage will be a constant. And we get a minimum $n_{1}$ satisfying the above requirement. We then want to get a contradiction to solve the problem. \\
We get a input $x$ with length $n_{1}$ that is accepted by $M$. And because that the total configuration sequences is less than $n_{1}$ / 3, we can easily use the drawer principle to get that there exists $i$,$j$,$k$ that have the same configuration sequence. Let $x$ be $a_{1}aa_{2}aa_{3}aa_{4}$  
in a format $a_{1}aa_{2}aa_{3}$. where $a$ is the symbol at $i$, $j$, $k$ and $a_{i}$, $i$ = 1,2,3,4 is a sequence. We construct $x_{1}$,$a_{1}aa_{3}aa_{4}$, we will compare the execution of $x$ and $x_{1}$ in $M$: \\
$M$ simulates same in $a{1}a$ until it is at position $i$ and heads right, and what $M$ simulates in $x_{1}$ in $i$ + 1 position is same as $M$ simulates
$x$ in $j$ + 1 position with same configuration sequence number $s_{1}$, and then when it hits position $k$ and heads right, $M$ will simulate same for two strings, and then when it is at position $k$ and heads left, it is with a configuration $s_{2}$, and loops as same like above, we can find for position $i$,$k$, it goes for the same configuration sequence for $x$ and $x_{1}$ and $M$ accepts $x_{1}$ too. \\
Similarly, we can construct $x_{2}$: $a_{1}aa_{2}aa_{4}$ and find that $M$ accepts $x_{2}$ too and $x_{2}$ share the same configuration as $x$ in position $i$, $j$. Now we will see the space usage for $x$, $x_{1}$, $x_{2}$, if $x$ reaches the maximum space usage $s(x)$ in the area, $a_{1}a$, $a_{3}aa_{4}$, 
because of the same configuration sequence in $i$, $k$, we can find $x_{1}$ will at a time reache a space usage $s(x)$, so $s(x_{1})$ $\geq$ $s(x)$. And if $x$ reaches the maximum space usage $s(x)$ in $a_{2}a$, because of the same configuration in position $j$ of $x$ and $x_{2}$, we can find $x_{2}$ will at a time reaches a space usage $s(x)$, so $s(x_{2})$ $\geq$ $s(x)$. So we can know there is a shorter sequence that uses $\geq$ $s(x)$ space. So we get a contradiction and solve this problem. \\
To prove the language $L$: $b(1)$\#$b(2)$\#$b(3)$...\#$b(k)$ is decided by a turning machine requires space $O(loglog(n))$, we just prove it starts at 000..0 and ends at 111...1 and each string is one-increment of last string, to check this, we need a counter to compare each digit of $i$ and $i + 1$, the lenght of binary representation of $i$ is $O(log(n))$, so the counter just costs the space usage $O(log(log(n))$. 
\part{}
\begin{enumerate}
\item Prove PSACE
We can use a recursion algorithm to solve this problem. \\
Because the graph won't change, just c and m position change \\
We construct two functions as below: \\
WinCat(c, m): \\
\itab{} \tab{set visit[c] = true} \\
\itab{} \tab{for every adjacent node i c has not visited:(return false if no such node)}\\
\itab{} \tab{} \tab{if i == m:}\\
\itab{} \tab{} \tab{} \tab{return true (catch the mouse}\\
\itab{} \tab{} \tab{else if NoloseMouse(i, m) is false:} \\ 
\itab{} \tab{} \tab{} \tab{return true} \\
\itab{} \tab{set visit[c] = false}\\
\itab{} \tab{return false}

NoloseMouse(c, m): \\
\itab{} \tab{if m == h:}\\
\itab{} \tab{} \tab{return true} \\
\itab{} \tab{for every adjacent node i of m which is not c: (return false if no such node)}\\
\itab{} \tab{} \tab{if i == h:} \\
\itab{} \tab{} \tab{} \tab{return true}\\ 
\itab{} \tab{} \tab{if WinCat(c, i) is false:}\\
\itab{} \tab{} \tab{} \tab{return true} \\
\itab{} \tab{return false(no such winning step)}\\
And HAPPY-CAT accepts input iff WinCat(c, m) is true, and because of the stack depth is at most $n$ which is the number of the vertexs. \\
And in each recursion we just use $O(log(n))$ space, so the total space usage is P, so the problem belongs to PSPACE.\\
\item Prove P 
We can find that if there exists a cycle of $G$, the mouse won't lose, at most draw, because the mouse just needs to run along the cycle, and the cat 
can not catch the mouse. \\
So we can use the algorithm dfs with visit flag to judge if a undirected graph has a cycle or not in time $O(n)$. \\
Then if there is a cycle, this graph won't be accepted by HAPPY-CAT unless c = m at the initial time. \\
If it has no cycle, then we use dijistra algorithm to calculate the shortest path with origin $h$, \\
if path[h][c] >= path[h][m]:\\
\itab{} \tab{return true}\\
\itab{} \tab{(because cat can go to the $h$ first, and because it is a tree, the mouse must be caught by the cat)}\\
else:\\
\itab{} \tab{return false (because the mouse can get to the hole first)} \\
We can see this algorithm is obviously in P. (cycle judge in P, shortest path algorithm in P)
\end{enumerate}
\end{document}
